\documentclass[10pt]{article}

\usepackage{amsfonts, amsthm, amsmath, fullpage, tikz, wrapfig, enumerate}

\newcommand{\card}[1]{\left| #1 \right|}
\newcommand{\nat}{\mathbb{N}}
\newcommand{\ints}{\mathbb{Z}}
\newcommand{\reals}{\mathbb{R}}
\newcommand{\chtitle}[1]{\noindent \vspace{5mm}\textbf{Chapter #1}\vspace{3mm}}
\newcommand{\images}{/home/gparker/classes/341/images}

\begin{document}
\begin{center}
\textbf{
CS 341 Automata Theory \\
Elaine Rich \\
Homework 9 \\
Due: Thursday, March 22}\\
\end{center}
\noindent
This assignment covers Chapter 13. \\

\begin{enumerate}[1)]

% ---
% 1
% ---

\item
For each of the following languages $L$, state whether $L$ is regular, context-free but not regular, or not context-free and prove your answer.
\begin{enumerate}[a)]
%a)
\item
$\{(ab)^na^nb^n:\ n > 0\}$.

%b)
\item
$\{xwx^R:\ x, w \in \{0, 1\}^+\}$.

%c)
\item
\{$a^ib^n$: i, $n > 0$ and $i = n$ or $i = 2n$\}.

%d)
\item
\{$0^i1^j$: $i, j \geq 0$ and $j = i^2$\}.

%e)
\item
$\{a^nb^mc^k:\ m \leq min(n, k)\}$.

%f)
\item
\{$x\#y$ : $x, y \in \{0, 1\}^*$ and when $x$ and $y$ are viewed as binary numbers, $y = x^2$\}.  For example, the string 100\#10000 $\in L$.
\end{enumerate}

%---
% 2
%---

\item
Give an example of a context-free language $L(\neq \Sigma ^*)$ that contains a subset $L_1$ that is not context-free.  Prove that $L$ is context free.  Describe $L_1$ and prove that it is not context-free.

% ---
% 3
% ---

\item
*  Give an example of a context-free language  $L$, other than one of the ones in the book, where  $\lnot L$ is not context-free.


% ---
% 4
% ---

\item
Are the context-free languages closed under each of the following operations?  Prove your answer.
\begin{enumerate}[a)]
%a
\item
$chop(L) = \{w:\ \exists x \in L\ (x = x_1cx_2 \land x_1 \in \Sigma _L^* \land x_2 \in \Sigma _L^* \land c \in \Sigma _L \land \card{x_1} = \card{x_2} \land w = x_1x_2)\}$.

%b
\item
Letter substitution.
\end{enumerate}

% ---
% 5
% ---

\item
Let $alt(L) = \{x:\ \exists y, n\ (y \in L, \card{y} = n, n > 0, y = a_1 \ldots a_n, \forall i \leq n (a_i \in \Sigma)$, and $x = a_1a_3a_5\ldots a_k$, where $k = ($if $n$ is even then $n-1$ else $n$))\}.
\begin{enumerate}[a)]
%a
\item
Consider $L = a^nb^n$.  Clearly describe $L_1 = alt(L)$.

%b
\item
Are the context free languages closed under the function alt?  Prove your answer.
\end{enumerate}

% ---
% 6
% ---

\item
Suppose that $L$ is context-free and $R$ is regular.  Is $R - L$ necessarily context-free?  Prove your answer.
\end{enumerate}
\end{document}
