\documentclass[10pt]{article}

\usepackage{amsfonts, amsthm, fullpage, tikz, wrapfig, enumerate}

\newcommand{\card}[1]{\left| #1 \right|}
\newcommand{\nat}{\mathbb{N}}
\newcommand{\ints}{\mathbb{Z}}
\newcommand{\reals}{\mathbb{R}}
\newcommand{\chtitle}[1]{\noindent \vspace{5mm}\textbf{Chapter #1}\vspace{3mm}}
\newcommand{\images}{/home/gparker/classes/341/images}

\begin{document}
\begin{center}
\textbf{
CS 341 Automata Theory \\
Elaine Rich \\
Homework 5 \\
Due Tuesday, February 14}\\
\end{center}
\noindent
This assignment covers Chapter 8. \\

\noindent
\textbf{Chapter 8}
\begin{enumerate}[1)]
% ---
% 1
% ---

\item
For each of the following languages $L$, state whether or not $L$ is regular.  Prove your answer:
\begin{enumerate}[a)]
%a
\item
$\{a^ib^j$ : $i, j \geq 0$ and $i + j = 5\}$.

%b
\item
$\{a^ib^j$ : $i, j \geq 0$ and $i - j = 5\}$.

%c
\item
$\{w = xy$ : $x, y \in \{a, b\}^*$ and $\card{x} = \card{y}$ and $\#_a(x) \geq \#_a(y)\}$.

%d
\item
$\{w = xyzy^Rx$ : $x, y, z \in \{a, b\}^*\}$.

%e
\item
$\{w = xyzy$ : $x, y, z \in \{0, 1\}^*\}$.

%f
\item
$\{w \in \{0, 1\}^* : \#_0(w) \neq \#_1(w)\}$.

%g
\item
* $\{w \in \{a, b\}^* : \exists x \in \{a, b\}^+\ (w = xx^Rx)\}$

%h
\item
$\{w \in \{a, b\}^*$ : the number of occurrences of the substring \texttt{ab} equals the number of occurrences of the substring \texttt{ba}\}.

%i
\item
* $\{w: w \in \{a – z\}^*$ and the letters of $w$ appear in reverse alphabetical order\}.  For example, \texttt{spoonfeed} $\in L$.

%j
\item
$L_0^*$, where $L_0 = \{ba^ib^ja^k,\ j \geq 0,\ 0 \leq i \leq k\}$.
\end{enumerate}

% ---
% 2
% ---

\item
For each of the following languages $L$, state whether $L$ is regular or not and prove your answer:
\begin{enumerate}[a)]
%a
\item
$\{uww^Rv$ : $u, v, w \in \{a, b\}^+\}$.

%b
\item
$\{xyzy^Rx\ :\ x, y, z \in \{a, b\}^+\}$.
\end{enumerate}


% ---
% 3
% ---

\item
Use the Pumping Theorem to complete the proof, given in Chapter L.3.1, that English isn’t regular.

% ---
% 4
% ---

\item
Prove by construction that the regular languages are closed under:

\begin{enumerate}[a)]
%a
\item
intersection.
 
%b
\item
* set difference.
\end{enumerate}

% --
% 5
% --
\item
Prove that the regular languages are closed under each of the following operations:
\begin{enumerate}[a)]
%a
\item
$pref(L) = \{w: \exists x \in \Sigma ^*\ (wx \in L)\}$.

%b
\item
$suff(L) = \{w: \exists x \in \Sigma ^*\ (xw \in L)\}$.
\end{enumerate}


% ---
% 6
% ---

\item
Using the definitions of  maxstring and  mix given in Section  8.6, give a precise definition of each of the 
following languages:

\begin{enumerate}[a)]
%a
\item
$maxstring(A^nB^n)$.

%b
\item
$maxstring(a^ib^jc^k, 1 \leq k \leq j \leq i)$.

%c
\item
$maxstring(L_1L_2)$, where $L_1 = \{w \in \{a, b\}^*$ : $w$ contains exactly one $a$\} and $L_2 = \{a\}$.

%d
\item
$mix((aba)^*)$.

%e
\item
$mix(a^*b^*)$.
\end{enumerate}

% ---
% 7
% ---

\item
Define the function $midchar(L) = \{c : \exists w \in L\ (w = ycz, c \in \Sigma _L, y \in \Sigma _L^*, z \in \Sigma _L^*,\ \card{y} = \card{z})\}$.  Answer each of the following questions and prove your answer:

\begin{enumerate}[a)]
%a
\item
Are the regular languages closed under $midchar$?  Prove your answer.

%b
\item
Are the nonregular languages closed under $midchar$?  Prove your answer.
\end{enumerate}

% ---
% 8
% ---
\item
Define the function $shuffle(L) = \{w : \exists x \in L$ ($w$ is some permutation of $x$)\}.  For example, if $L = \{ab,\ abc\}$, then $shuffle(L) = \{ab,\ abc,\ ba,\ acb,\ bac,\ bca,\ cab,\ cba\}$.  Are the regular languages closed under $shuffle$?  Prove your answer.

% ---
% 9
% ---
\item
Consider any function $f(L_1) = L_2$, where $L_1$ and $L_2$ are both languages over the alphabet $\Sigma = \{0, 1\}$.  We say that function $f$ is \textit{nice} iff  $L_2$ is regular iff $L_1$ is regular.  For each of the following functions, state whether or not it is nice and prove your answer:

\begin{enumerate}[a)]
%a
\item
$f(L) = L^R$.

%b
\item
$f(L)  =  \{w:\ w$ is formed by taking a string in $L$ and replacing all  \texttt{1}'s with  \texttt{0}'s and leaving the  \texttt{0}'s unchanged\}.

%c
\item
$f(L) = L \cup 0^*$

%d
\item
$f(L)  =  \{w:\ w$ is formed by taking a string in $L$ and replacing all \texttt{1}'s with  \texttt{0}'s and all  \texttt{0}'s with  \texttt{1}'s (simultaneously)\}.

%e
\item
$f(L) = \{w: \exists x \in L\ (w = x\texttt{00})\}$.
\end{enumerate}

% ---
% 10
% ---

\item
Consider the language $L = \{x\texttt{0}^ny\texttt{1}^nz\ :\ n \geq 0, x \in P, y \in Q, z \in R\}$, where $P$, $Q$, and $R$ are nonempty sets over the alphabet $\{0, 1\}$.  Can you find regular languages $P$, $Q$, and $R$ such that $L$ is not regular?  Can you find regular languages $P$, $Q$, and $R$ such that $L$ is regular?
\end{enumerate}
\end{document}
